Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-01-22 doc/invt.tex
add predict.c
add errsbin.gp

@@

The total error of the bias potential can be written in
terms of $x_i$,
\begin{equation}
\mathcal E
=
\sum_{i = 1}^n p_i \, \langle x_i^2 \rangle
=
\sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle,
\label{eq:error_sum}
\end{equation}
where
$\bar v = \sum_{i = 1}^n p_i v_i$.
%
Our aim is to find the $\alpha(t)$
that minimizes Eq. \eqref{eq:error_sum}.
%
Fortunately, we shall find that it suffices to consider
a simpler one-variable problem
in this case.

@@

The error function can be obtained
from Eq. \eqref{eq:eig_orthonormal_cols}
as
\begin{align}
  \mathcal E
  &\equiv
  \sum_{i = 1}^n p_i \, x_i^2
  =
  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \, x_i \, x_j
  =
  \sum_{k = 1}^n y_k^2.
  \label{eq:y2_sum}
\end{align}
%
Note that there is always one mode, say
$y_1 = \sum_{j=1}^n p_j \, x_j$
(corresponding to $\lambda_1 = 1$),
vanishes identically by Eq. \eqref{eq:px_sum}.
%
Thus, we can start the sum in Eq. \eqref{eq:y2_sum}
from $k = 2$.

In analogous to Eq. \eqref{eq:x2t_average},
we can rewrite Eq. \eqref{eq:y2_sum} as
%
\begin{align}
  \mathcal E
  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
  %\notag
  %\\
  &=
  \int_0^t \int_0^t
  \sum_{i = 2}^n
  \dot u_i\bigl( q(t') \bigr) \,
  \dot u_i\bigl( q(t'') \bigr) \,
  \kappa_i(t' - t'') \, dt' \, dt'',
  \label{eq:error_mbin}
\end{align}
%
where
\begin{align*}
  u_i(z)
  &\equiv
  e^{\lambda_i \, [q(t') - z]},
  \\
  \kappa_i(t - t')
  &\equiv
  \left\langle
    \eta_i(t) \, \eta_i(t')
  \right\rangle.
\end{align*}
%
\note{The extremal condition of $\mathcal E$ is given by
%
\begin{align}
\sum_{i=2}^n
\lambda_i \, u_i\bigl( q(\tau) \bigr)
\int_0^t
\ddot u_i\bigl( q(t') \bigr) \, \kappa_i(t' - \tau) \, dt' = 0.
\label{eq:optimal_mbin}
\end{align}
%
If all $\lambda_i$, hence all $u_i$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
Unfortunately, the above expression
appears to be useless in a general setting.
}

@@

In the asymptotic regime,
we can approximate $\kappa_i(t)$ as
%
\begin{equation}
  \kappa_i(t) \approx \Gamma_i \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_i = \sum_{t = -\infty}^\infty \kappa_i(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_mbin} is simplified as
%
\begin{align}
  \mathcal E
  &=
  \int_0^t
  \sum_{i = 2}^n
  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
  \label{eq:error_mbin1}
\end{align}
%
The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\mathcal E$
that is about $\alpha_{\max}^2$ times as large.

\note{This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_i(\omega) = \Gamma_i
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
  $$
  which leads to a correction to $\mathcal E$
  $$
  \Gamma^{(2)}_i
  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
  \propto
  \alpha^2 \, \mathcal E.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}

@@

We now show
that the single-bin scheme
is an optimal scheme.
%
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence same $\Gamma_i$'s,
but with different eigenvalues, $\lambda_i$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of $\beta_i \ge 0$
%
\begin{align}
&
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr)
\right)
%
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i^2
\right)
%
\notag
\\
&
\ge
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i \, \dot u_i\bigl( q(t') \bigr)
\right)^2
\notag
\\
&
=
\left(
  \sum_{i = 2}^n \Gamma_k \, c_i
    \left[
      1 - e^{ -\lambda_i q(t) }
    \right]
\right)^2.
\notag
\end{align}
%
\note{A short proof of the above inequality
  is the following.
  %
  First, we have
$$
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \,
    \left( \dot u_i \, X - c_i \right)^2
\ge 0,
$$
  for any $X$.
  %
  Thus, the left-hand side,
  expanded as a quadratic polynomial of $X$,
  $A \, X^2 + B \, X + C$,
  must have a non-positive discriminant,
  $B^2 - 4 \, A \, C$.
  %
  This condition gives the desired inequality.
  \\%
}
%
Note that the last expression depends on
$q(t')$ only through the end point value,
which is fixed in the variation process.
%
Thus, the inequality sets a lower bound
of the error $\mathcal E$
expressed in Eq. \eqref{eq:error_mbin1},
%
The equality is achieved
if $\dot u_i\left( q(t') \right) = c_i$
for all $i = 2, \dots, n$ at any $t'$,
up to a multiplicative factor.
%
Solving this set of equations yields
$$
\alpha(t') = \frac{              1              }
                  { \lambda_i \, (t' + t_{i0} ) },
\quad
\mathrm{with\;\;}
t_{i0} = \frac{             t            }
              { e^{ \lambda_i q(t) } - 1 }.
$$
Such a solution is possible only if
\begin{equation}
\lambda_2 = \dots = \lambda_n.
\end{equation}
%
\note{
  We add some steps for the above solution.
  Integrating $\dot u_i = c_i$ yields
  \begin{equation}
  u_i(t') = c_i \left(t' + t_{i0} \right).
  \label{eq:ui_solution}
  \end{equation}
  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
  we have
  $$
  c_i = \frac{ 1 }{ t + t_{i0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{i0} } { t + t_{i0} }
  =
  e^{ -\lambda_k \, q(t) }.
  $$
  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
  $$
  -\lambda_i \, q(t) + \lambda_i \, q(t')
  = \log c_i + \log\left( t' + t_{i0} \right).
  $$
  Differentiating this with respect to $t'$ yields
  $$
  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
  $$
}

@@

If we wish to find,
for a given updating matrix $\mathbf w$,
a near-optimal $\alpha(t)$,
%
then a good candidate is
%
\begin{equation}
\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
\label{eq:alpha_trial}
\end{equation}
%
for some optimal value of $\lambda$.
%
Using Eq. \eqref{eq:alpha_trial}
in Eq. \eqref{eq:error_mbin1} yields
%
\begin{align}
\mathcal E
&=
\frac{1}{t+t_0}
\sum_{i=2}^n
  \frac{       \Gamma_i \, \lambda_i^2      }
       { \lambda \, (2 \lambda_i - \lambda) }
\left[
  1 - \left(
        1 + \frac{ t }{ t_0 }
      \right)^{1 - 2 \, \lambda_i/\lambda}
\right]
\notag
\\
&\stackrel{      t \gg t_0      }
          { =\joinrel=\joinrel= }
%
\sum_{i = 2}^n
\begin{dcases}
  \frac{ 1 }
       { t }
  \frac{         \Gamma_i \, \lambda_i^2       }
       { \lambda \, (2 \, \lambda_i - \lambda) }
  &
  \mathrm{if \;} \lambda < 2 \, \lambda_i,
  \\%[1em]
  %
  %
  \frac{ \Gamma_i }
       {  4 \, t  }
  \ln \frac{ t + t_0 }
           {  t_0    }
  &
  \mathrm{if \;} \lambda = 2 \, \lambda_i,
  \\%[1em]
  %
  %
  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1 } }
       {   t^{ 2 \, \lambda_i / \lambda      } }
  \frac{      \Gamma_i \, \lambda_i^2          }
       { \lambda \, (\lambda - 2 \, \lambda_i) }
  &
  \mathrm{if \;} \lambda > 2 \, \lambda_i.
\end{dcases}
\label{eq:y2_sum_perfect_invt}
\end{align}
%
\note{In the above expression,
$t + t_0$ are replaced by $t$.
%
Adding back $t_0$, we get
$$
\begin{aligned}
  \mathcal E
  \stackrel{         t \to \infty         }
           { =\joinrel=\joinrel=\joinrel= }
  \sum_{i = 2}^n
  \begin{dcases}
    \frac{    1    }
         { t + t_0 }
    \frac{       \Gamma_i \, \lambda_i^2         }
         { \lambda \, (2 \, \lambda_i - \lambda) }
    &
    \mathrm{if \;} \lambda < 2 \, \lambda_i,
    \\%[1em]
    %
    %
    \frac{    \Gamma_i    }
         { 4 \, (t + t_0) }
    \ln \frac{ t + t_0 }
             {   t_0  }
    &
    \mathrm{if \;} \lambda = 2 \, \lambda_i,
    \\
    %
    %
    \frac{  t_0^{ 2 \, \lambda_i / \lambda  - 1}  }
         { (t + t_0)^{ 2 \, \lambda_i / \lambda } }
    \frac{      \Gamma_i \, \lambda_i^2          }
         { \lambda \, (\lambda - 2 \, \lambda_i) }
    &
    \mathrm{if \;} \lambda > 2 \, \lambda_i.
  \end{dcases}
\end{aligned}
$$
}
%
The last two cases, $\lambda \ge 2\lambda_i$,
are asymptotically suboptimal
as they decay slower than $1/t$.
%
Thus, the optimal $\lambda$
satisfies $\lambda < 2 \min \lambda_i$,
and in this case,
we can seek that
the optimal value from
the smallest real root of
%
\begin{equation}
\sum_{i = 2}^n
\frac{ \lambda_i - \lambda }
{ \left(2 - \lambda/ \lambda_i \right)^2 }
= 0.
\label{eq:optimal_lambda_approx}
\end{equation}

@@

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral.
%
Particularly,
for perfect sampling,
we get from Eq. \eqref{eq:Gamma_perfect},
$$
\begin{aligned}
\mathcal E
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
  %
$$
\begin{aligned}
\mathcal E
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
  \mathcal E
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\mathcal E
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.

>>>>

\newcommand{\Err}{E}

@@

The total error of the bias potential can be written in
terms of $x_i$,
\begin{equation}
\Err
=
\sum_{i = 1}^n p_i \, \langle x_i^2 \rangle
=
\sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle,
\label{eq:error_sum}
\end{equation}
where
$\bar v = \sum_{i = 1}^n p_i v_i$.
%
Our aim is to find the $\alpha(t)$
that minimizes Eq. \eqref{eq:error_sum}.
%
Fortunately, we shall find that it suffices to consider
a simpler one-variable problem
in this case.

@@

The error function can be obtained
from Eq. \eqref{eq:eig_orthonormal_cols}
as
\begin{align}
  \Err
  &\equiv
  \sum_{i = 1}^n p_i \, x_i^2
  =
  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \, x_i \, x_j
  =
  \sum_{k = 1}^n y_k^2.
  \label{eq:y2_sum}
\end{align}
%
Note that there is always one mode, say
$y_1 = \sum_{j=1}^n p_j \, x_j$
(corresponding to $\lambda_1 = 1$),
vanishes identically by Eq. \eqref{eq:px_sum}.
%
Thus, we can start the sum in Eq. \eqref{eq:y2_sum}
from $k = 2$.

In analogous to Eq. \eqref{eq:x2t_average},
we can rewrite Eq. \eqref{eq:y2_sum} as
%
\begin{align}
  \Err
  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
  %\notag
  %\\
  &=
  \int_0^t \int_0^t
  \sum_{i = 2}^n
  \dot u_i\bigl( q(t') \bigr) \,
  \dot u_i\bigl( q(t'') \bigr) \,
  \kappa_i(t' - t'') \, dt' \, dt'',
  \label{eq:error_mbin}
\end{align}
%
where
\begin{align*}
  u_i(z)
  &\equiv
  e^{\lambda_i \, [z - q(t)]},
  \\
  \kappa_i(t - t')
  &\equiv
  \left\langle
    \eta_i(t) \, \eta_i(t')
  \right\rangle.
\end{align*}
%
\note{The extremal condition of $\Err$ is given by
%
\begin{align}
\sum_{i=2}^n
\lambda_i \, u_i\bigl( q(\tau) \bigr)
\int_0^t
\ddot u_i\bigl( q(t') \bigr) \, \kappa_i(t' - \tau) \, dt' = 0.
\label{eq:optimal_mbin}
\end{align}
%
If all $\lambda_i$, hence all $u_i$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
Unfortunately, the above expression
appears to be useless in a general setting.
}

@@

In the asymptotic regime,
we can approximate $\kappa_i(t)$ as
%
\begin{equation}
  \kappa_i(t) \approx \Gamma_i \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_i = \sum_{t = -\infty}^\infty \kappa_i(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_mbin} is simplified as
%
\begin{align}
  \Err
  &=
  \int_0^t
  \sum_{i = 2}^n
  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
  \label{eq:error_mbin1}
\end{align}
%
The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\Err$
that is about $\alpha_{\max}^2$ times as large.

\note{This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_i(\omega) = \Gamma_i
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
  $$
  which leads to a correction to $\Err$
  $$
  \Gamma^{(2)}_i
  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
  \propto
  \alpha^2 \, \Err.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}

@@

We now show
that the single-bin scheme
is an optimal scheme.
%
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence same $\Gamma_i$'s,
but with different eigenvalues, $\lambda_i$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of $\beta_i \ge 0$
%
\begin{align}
&
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr)
\right)
%
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i^2
\right)
%
\notag
\\
&
\ge
\left(
  \int_0^t dt'
    \sum_{i = 2}^n
      \Gamma_i \, c_i \, \dot u_i\bigl( q(t') \bigr)
\right)^2
\notag
\\
&
=
\left(
  \sum_{i = 2}^n \Gamma_k \, c_i
    \left[
      1 - e^{ -\lambda_i q(t) }
    \right]
\right)^2.
\notag
\end{align}
%
\note{A short proof of the above inequality
  is the following.
  %
  First, we have
$$
\int_0^t
  dt' \sum_{i = 2}^n \Gamma_i \,
    \left( \dot u_i \, X - c_i \right)^2
\ge 0,
$$
  for any $X$.
  %
  Thus, the left-hand side,
  expanded as a quadratic polynomial of $X$,
  $A \, X^2 + B \, X + C$,
  must have a non-positive discriminant,
  $B^2 - 4 \, A \, C$.
  %
  This condition gives the desired inequality.
  \\%
}
%
Note that the last expression depends on
$q(t')$ only through the end point value,
which is fixed in the variation process.
%
Thus, the inequality sets a lower bound
of the error $\Err$
expressed in Eq. \eqref{eq:error_mbin1},
%
The equality is achieved
if $\dot u_i\left( q(t') \right) = c_i$
for all $i = 2, \dots, n$ at any $t'$,
up to a multiplicative factor.
%
Solving this set of equations yields
$$
\alpha(t') = \frac{              1              }
                  { \lambda_i \, (t' + t_{i0} ) },
\quad
\mathrm{with\;\;}
t_{i0} = \frac{             t            }
              { e^{ \lambda_i q(t) } - 1 }.
$$
Such a solution is possible only if
\begin{equation}
\lambda_2 = \dots = \lambda_n.
\end{equation}
%
\note{
  We add some steps for the above solution.
  Integrating $\dot u_i = c_i$ yields
  \begin{equation}
  u_i(t') = c_i \left(t' + t_{i0} \right).
  \label{eq:ui_solution}
  \end{equation}
  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
  we have
  $$
  c_i = \frac{ 1 }{ t + t_{i0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{i0} } { t + t_{i0} }
  =
  e^{ -\lambda_k \, q(t) }.
  $$
  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
  $$
  -\lambda_i \, q(t) + \lambda_i \, q(t')
  = \log c_i + \log\left( t' + t_{i0} \right).
  $$
  Differentiating this with respect to $t'$ yields
  $$
  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
  $$
}

@@

If we wish to find,
for a given updating matrix $\mathbf w$,
a near-optimal $\alpha(t)$,
%
then a good candidate is
%
\begin{equation}
\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
\label{eq:alpha_trial}
\end{equation}
%
for some optimal value of $\lambda$.
%
Using Eq. \eqref{eq:alpha_trial}
in Eq. \eqref{eq:error_mbin1} yields
%
\begin{align}
\Err
&=
\frac{1}{t+t_0}
\sum_{i=2}^n
  \frac{       \Gamma_i \, \lambda_i^2      }
       { \lambda \, (2 \lambda_i - \lambda) }
\left[
  1 - \left(
        1 + \frac{ t }{ t_0 }
      \right)^{1 - 2 \, \lambda_i/\lambda}
\right]
\notag
\\
&\stackrel{      t \gg t_0      }
          { =\joinrel=\joinrel= }
%
\sum_{i = 2}^n
\begin{dcases}
  \frac{ 1 }
       { t }
  \frac{         \Gamma_i \, \lambda_i^2       }
       { \lambda \, (2 \, \lambda_i - \lambda) }
  &
  \mathrm{if \;} \lambda < 2 \, \lambda_i,
  \\%[1em]
  %
  %
  \frac{ \Gamma_i }
       {  4 \, t  }
  \ln \frac{ t + t_0 }
           {  t_0    }
  &
  \mathrm{if \;} \lambda = 2 \, \lambda_i,
  \\%[1em]
  %
  %
  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1 } }
       {   t^{ 2 \, \lambda_i / \lambda      } }
  \frac{      \Gamma_i \, \lambda_i^2          }
       { \lambda \, (\lambda - 2 \, \lambda_i) }
  &
  \mathrm{if \;} \lambda > 2 \, \lambda_i.
\end{dcases}
\label{eq:y2_sum_perfect_invt}
\end{align}
%
\note{In the above expression,
$t + t_0$ are replaced by $t$.
%
Adding back $t_0$, we get
$$
\begin{aligned}
  \Err
  \stackrel{         t \to \infty         }
           { =\joinrel=\joinrel=\joinrel= }
  \sum_{i = 2}^n
  \begin{dcases}
    \frac{    1    }
         { t + t_0 }
    \frac{       \Gamma_i \, \lambda_i^2         }
         { \lambda \, (2 \, \lambda_i - \lambda) }
    &
    \mathrm{if \;} \lambda < 2 \, \lambda_i,
    \\%[1em]
    %
    %
    \frac{    \Gamma_i    }
         { 4 \, (t + t_0) }
    \ln \frac{ t + t_0 }
             {   t_0  }
    &
    \mathrm{if \;} \lambda = 2 \, \lambda_i,
    \\
    %
    %
    \frac{  t_0^{ 2 \, \lambda_i / \lambda  - 1}  }
         { (t + t_0)^{ 2 \, \lambda_i / \lambda } }
    \frac{      \Gamma_i \, \lambda_i^2          }
         { \lambda \, (\lambda - 2 \, \lambda_i) }
    &
    \mathrm{if \;} \lambda > 2 \, \lambda_i.
  \end{dcases}
\end{aligned}
$$
}
%
The last two cases, $\lambda \ge 2\lambda_i$,
are asymptotically suboptimal
as they decay slower than $1/t$.
%
Thus, the optimal $\lambda$
satisfies $\lambda < 2 \min \lambda_i$,
and in this case,
we can seek that
the optimal value from
the smallest real root of
%
\begin{equation}
\sum_{i = 2}^n
\frac{ \lambda_i - \lambda }
{ \left(2 - \lambda/ \lambda_i \right)^2 }
= 0.
\label{eq:optimal_lambda_approx}
\end{equation}

@@

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral.
%
Particularly,
for perfect sampling,
we get from Eq. \eqref{eq:Gamma_perfect},
$$
\begin{aligned}
\Err
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
  %
$$
\begin{aligned}
\Err
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
\Err
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\Err
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.
